---
title: "Untitled"
author: "Connor Concannon"
date: "September 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('e1071')
options(scipen=1000)

library(DMwR)
library(ggplot2)
library(dplyr)
library(DescTools)
library(caret)
library(stringr)
library(lubridate)

c <- read.csv('Crashes.csv')
str(c);names(c)
#Desc(c,plotit=T)
Desc(c$NUMBER.OF.PERSONS.KILLED)

#Recode, dummy
killed <- ifelse(c$NUMBER.OF.PERSONS.KILLED>0,1,0)
times <- c %>% mutate(yr=year(DATE),
                   mo=month(DATE),
                   da=day(DATE),
                   hr=hour(hm(as.character(TIME))))

boro <- dummyVars(~BOROUGH,data=c)
factor1 <- dummyVars(~CONTRIBUTING.FACTOR.VEHICLE.1,data=c)
vehicle <- dummyVars(~VEHICLE.TYPE.CODE.1 ,data=c)
c$ZIP.CODE <- as.character(c$ZIP.CODE)

zip <- dummyVars(~ZIP.CODE,data=c)

boro <-  data.frame(predict(boro,c))
factor1 <- data.frame(predict(factor1,c))
zip <- data.frame(predict(zip,c))
vehicle <- data.frame(predict(vehicle,c))

#bind
c2 <- cbind(c,boro)
c2 <- cbind(c2,factor1)
c2 <- cbind(c2,times)
c2 <- cbind(c2,killed)
#c2 <- cbind(c2,zip)
c2 <- cbind(c2,vehicle)

names(c2)

c2 <- c2[,-c(1:16,71:86)]

c2$killed <- factor(ifelse(c2$killed==1,'Yes','No'))


c2 <- na.omit(c2)

glimpse(c2)
names(c2)
#Split
c3 <- c2[sample(nrow(c2),10000),]
c3 <- c2
trainindex <- createDataPartition(c3$killed,p=.8,list=F)

train <- c3[trainindex,]
test <- c3[-trainindex,]


#Model
form <- as.formula(killed~.)


ctrl <- trainControl(method='repeatedcv',number=5,repeats=5,classProbs = T)

#rf <- train(form,data=train,method='rf',trControl=ctrl,verbose=T,mtry=10)
#gbm <- train(form,data=train,method='gbm',trControl=ctrl,verbose=T)
#gbm
#plot(gbm)
#summary(gbm)



```

```{r}

smote <- SMOTE(form,data=train)
glm <- train(form,data=train,method='glm')
table(smote$killed)


regular <- train(form,data=train,
                       method='treebag',
                       nbagg=50,
                       metric='ROC',
                       trControl=ctrl)

smote_outside <- train(form,data=smote,
                       method='treebag',
                       nbagg=50,
                       metric='ROC',
                       trControl=ctrl)

outside_models <- list(regular=regular,
                       smote=smote_outside)

outside_resample <- resamples(outside_models)

test_roc <- function(model,data){
  library(pROC)
  roc_obj <- roc(data$killed,
                 predict(model,data,type="prob")[, "Yes"],
                 levels=c("No","Yes"))
  ci(roc_obj)
}


outside_test <- lapply(outside_models,test_roc,data=test)
outside_test <- lapply(outside_test,as.vector)
outside_test <- do.call("rbind",outside_test)
colnames(outside_test) <- c('lower','ROC','upper')
outside_test <- as.data.frame(outside_test)

summary(outside_resample,metric="ROC")

pred <- predict(smote_outside,test,type='prob')

test <- cbind(test,pred)




glimpse(test)

qplot(test$killed)

varImp(smote_outside)

qplot(test$hr,fill=test$killed,binwidth=1)

detach('package:plyr',unload=T)
test %>% 
  group_by(da) %>% 
  dplyr::summarise(num=n(),
            Killed=sum(ifelse(killed=='Yes',1,0))) %>% 
  mutate(perc=Killed/num) %>% 
  ggplot(aes(x=da,y=perc))+geom_bar(stat='identity')


